---
title: "Data 607: Week 7 Assignment 3"
author: "Talha Muhammad"
date: "October 16, 2016"
output: html_document
---
       
## Overview 
Assignment 3 consists of developing some toy data on book publishers in JSON, XML and HTML. These data sets are then coverted to R data frames to show a simple case of how these sources of data might be handled. Below we proceed in the following order      
1.JSON     
2.XML     
3.HTML    

We first created the above datasets in the JSON, XML and HTML formats, "by hand". The respective formats were then imported into R using appopriate packages. 
     
```{r filepath, include=TRUE}
#file path for all three datasets
datadir=file.path("C://Users//talha//Documents//Training//CUNY Classes//IS607//Week7//")
```

### JSON data
```{r setup, warning=FALSE,message=FALSE}
#load packages in JSONLite
library(jsonlite)
filename=file.path(datadir,"bookdata.js")
```

```{r Analysis}
#read file names
bdata<-readLines(filename)
bdata
bdata_str<-paste(bdata,"")
#get JSON data
jsondata<-fromJSON(bdata_str)
#class and attributes of jsondata
class(jsondata)
length(jsondata)
names(jsondata)
lapply(jsondata,class)
lapply(jsondata,length)
lapply(jsondata,dim)
#view different data elements
head(jsondata[[1]],n=4)
head(jsondata[[2]],n=4)
#create data frame from JSON data
jsondf<-as.data.frame(jsondata)
#the data frame
jsondf
dim(jsondf)
```
      
### XML data   

```{r XML}
#load a the XML library
library(XML)
```

```{r}
#load the file
filename=file.path(datadir,"bookdata.xml")
#parse the XML file
xml_str<-xmlTreeParse(filename)
class(xml_str)
#extract the root node
root<-xmlRoot(xml_str)
# extract children node
child<-xmlChildren(root)
# extract subchildren node
schild<-xmlChildren(child[[1]])

# extract nodes using XPATH notation
getNodeSet(root,"//booksnum[@ID]")
getNodeSet(root,"//author")
getNodeSet(root,"//title")
getNodeSet(root,"//type")
getNodeSet(root,"//publisher")
getNodeSet(root,"//language")

sapply(schild,xmlName)
sapply(child,xmlValue)
sapply(child,xmlAttrs)
sapply(schild,xmlValue)
# develop data frame from XML data
author_df0<-as.data.frame(sapply(getNodeSet(root,"//booksnum[@ID]"),xmlAttrs))
author_df1<-as.data.frame(sapply(getNodeSet(root,"//author"),xmlValue))
author_df2<-as.data.frame(sapply(getNodeSet(root,"//title"),xmlValue))
author_df3<-as.data.frame(sapply(getNodeSet(root,"//type"),xmlValue))
author_df4<-as.data.frame(sapply(getNodeSet(root,"//publisher"),xmlValue))
author_df5<-as.data.frame(sapply(getNodeSet(root,"//language"),xmlValue))
author_df<-cbind(author_df0,author_df1,author_df2,author_df3,author_df4,author_df5)
#column names
colnames(author_df)<-c("ID",sapply(schild,xmlName))
#dataframe from XML data
author_df
```
     
### HTML data    
      
```{r HTML}
filename=file.path(datadir,"bookdata.html")
# parse the HTML file
html_str<-htmlParse(filename)
# read tables
tables<-readHTMLTable(html_str,stringsAsFactors=FALSE)
# develop data frames
html_df<-as.data.frame(tables[[1]])
html_df