---
title: "Project 4"
author: "Talha Muhammad"
date: "November 16, 2016"
output: html_document
---

### Overview
Migrate an SQL database to a No-SQL database. I will use the *flights* database as the SQL database and migrate to the NOSQL database *MongoDB*.

```{r setup, include=TRUE, warning=FALSE, message=FALSE}
#load Mongo DB, JSON and other packages
library(mongolite)
library(tidyr)
library(dplyr)
library(jsonlite)
library(RMySQL)
```

###Query a MongoDB Collection
For illustration I query the restaurants dataset using MongoDB. Data has been previously loaded into MongoDB:

```{r database}
# Query the Mongo database
m<-mongo(collection="restaurants", db="test")
# summary counts by borough
m$aggregate('[{"$group": { "_id": "$borough", "count": {"$sum": 1 } } }]')
# select the data
restman<-m$find('{ "borough": "Manhattan" }')
restman<-flatten(restman, recursive=TRUE)
# data header
head(restman)
#summary by cuisine in Manhattan
rest_cu<-restman %>% group_by(cuisine) %>% count(borough,sort=TRUE)
# summary of restuarants in the City
head(rest_cu)
```

### Query the SQL database
I query the flights database from the local SQL server. I query the planes table in the flights database.    

```{r flightsdata}
rmysql.settingsfile<-"C:/Program Files/MySQL/MySQL Server 5.0/my.ini"
con <- dbConnect(RMySQL::MySQL(), dbname = "flights", username="root", password="password")
flights.planes<-dbGetQuery(con,"SELECT * from planes")
dbDisconnect(con)
# data header
head(flights.planes)
```

###Import to MongoDB

I add the data SQL database **flights.planes** as a dataset in MongoDB. 

```{r export}
fp_js<-mongo(collection="flights", db="planes",url="mongodb://localhost")
# insert the database into a MongoDB
fp_js$insert(flights.planes)            
# select the data for manufacturer
planes_embraer<-fp_js$find('{"manufacturer":"EMBRAER"}')
#head of the data
head(planes_embraer)
```

