---
title: 'Data 605, Visualization: Hw 1'
author: "Talha Muhammad"
date: "September 9, 2017"
output: html_document
---

```{r setup, include=FALSE}
library(RCurl)
library(ggplot2)
library(dplyr)
incmag_data<-read.csv(text=getURL("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA608/master/lecture1/Data/inc5000_data.csv"),header=TRUE,sep=",")
```

## Home Work 1 Data 605
Develop visualizations using this dataset consisting of the 5000 fastest growing companies in the US.    

After loading the data we develop some summaries of the dataset. 
   
```{r summary}
glimpse(incmag_data)
summary(incmag_data)
head(incmag_data)
```
### Question 1

**Create a graph that shows the distribution of companies in the dataset by State**

```{r fig1}
# figure 1, show the number of 
p<-ggplot(count(incmag_data,State), aes(x=reorder(State,n),n))
p+geom_bar(stat="identity",fill="darkblue")+xlab("State")+ylab("# Firms")+ggtitle("Number of Firms by State")+coord_flip()+theme(axis.text.y = element_text(size=8))
#Following Tufte's guidance on minimalism
p<-ggplot(count(incmag_data,State), aes(x=reorder(State,n),n))
p+geom_bar(stat="identity",fill="darkblue",alpha=0.7)+xlab("State")+ylab("# Firms")+ggtitle("Number of Firms by State")+coord_flip()+theme_bw()+theme(axis.text.y = element_text(size=7))
```


Let's dig in on the State with the 3rd most companies in the data set:    

### Question 2   

**Create a plot of average employment by industry for companies in this state**   
only use cases with full data. Your graph should show how variable the ranges are,and exclude outliers.


```{r fig2}
#filter data to focus on the 3rd State
select_data<-filter(incmag_data,State=="NY")
#filter data to remove any missing values
select_data$complete<-complete.cases(select_data)
select_data<-filter(select_data,complete)
#filter-out the outliers
select_data<-select_data %>% group_by(Industry) 
select_data$rank<-percent_rank(select_data$Employees)
select_data<-filter(select_data,rank<0.975,rank>0.025)
#check the result
summary(select_data$rank)
##plot the data
p<-ggplot(select_data,aes(Industry,Employees,color=Industry))
#use the box plot and change some formatting
p1<-p+geom_boxplot()+guides(colour="none")+ theme(axis.text.x = element_text(angle=90,vjust=0.5))+scale_y_continuous(breaks=seq(0,1600,200))
#add means to the chart
p1+stat_summary(fun.y="mean",geom="point",shape=23,size=2,fill="white")
```

### Question 3 
**Create a chart showing revenue per employee**

```{r fig3}
#calculate revenue per emp
select_data<-mutate(select_data,rev_emp=(Revenue/Employees)/1000)
#group and calculate the mean revenue per employee by Industry
select_data<-select_data %>% group_by(Industry)
revenue_data<-summarize(select_data,mrev=mean(rev_emp))
#filter data to focus on the 3rd State
ggplot(revenue_data, aes(x=mrev,y=reorder(Industry,mrev))) + geom_point(color="red",size=2)+theme_minimal()+scale_x_continuous(breaks=seq(100,800,100))+xlab("Revenue Per Employee (in $ thousands)")+ggtitle("Mean Revenue per Employee by Industry")+ylab("Industry")
```
