---
title: "Data 607, Assignment Week5"
author: "Talha Muhammad"
date: "October 2, 2016"
output: html_document
---

```{r filepath, include=FALSE}
wkdir=file.path("C://Users//talha//Documents//Training//CUNY Classes//IS607//Week5")
setwd(wkdir)
```

## Overview    
       
The purpose of this assignment is to use the packages "tidyr" and "dplyr" and the different functions that are part of the packages. The exercises below reshape and use the data to illustrate the different packages. 


```{r setup, warning=FALSE,message=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(plotly)
```

```{r fileread}
#read in the file
numbersdata<-read.csv("numbersense.csv",header=TRUE,sep=",")
# this is the original data set we are working with
head(numbersdata)
# selectdata to remove #NA's
selectdata<-numbersdata %>%  
  filter(!is.na(Phoenix)) %>% 
    rename(Airline=X,Arrival=X.1,"Los Angeles"=Los.Angeles,"San Diego"=San.Diego,"San Francisco"=San.Francisco)

#replace missing values
selectdata$Airline<-as.character(selectdata$Airline)
selectdata$Airline[selectdata$Airline==""]<-c("Alaska","AM West")
# renamed dataset
selectdata

#create tidy dataset and sort the data
tidydata<-selectdata %>% 
  gather(key=region,value=Number,3:7)
tidydata<-arrange(tidydata,Airline,Arrival)
# the tidy dataset
head(tidydata)

# total flights delayed and and Ontime
tidydata %>% 
  group_by(Airline,Arrival) %>% 
  summarize(total=sum(Number))
# total flights by airline
tidydata %>% 
  group_by(Airline) %>% 
  summarize(total=sum(Number))

#reshape the data and calculate proportions
# use the mutate function to total and proportion of delayed flights
spreadata<-tidydata %>% 
  spread(Arrival,Number) %>% 
  mutate(total=delayed+`on time`,prop=delayed/total)
```
      
      
```{r plots}
#create plot
# plot of delayed airlines
ggplotly(ggplot(filter(tidydata,Arrival=="delayed"),aes(region,Number,color=Airline))+geom_point(size=3)+ggtitle("Delayed Flights")+ylab("Number of Delayed Flights"))
 

#bar plot of delayed flights
ggplotly(ggplot(spreadata,aes(x=region,y=prop))+geom_bar(aes(fill=Airline),stat="identity",position="dodge")+ylab("Proportion of Delayed Flights")+ggtitle("Delayed Flights by Region"))
```
       
        
The steps shown serve as an illustration of the way different functions within the "dplyr" and "tidyr" packages work and how they can simplify data analysis. 
   
   